The Ontological Modal Collapse \\ as a Collapse of the Square of Opposition

### 2016-03-25 Papers/2014/Universalis/ModalCollapse.tex

%----------classification, keywords, date
\subjclass{
Prim. 03A02;  % Philosophical aspects of logic and foundations
Sec. 68T02 % Artificial Intelligence
}

\keywords{Modal Logics, Higher-Order Logics, Ontological Argument}

@@

Nevertheless, G{\"o}del's work still leaves room for criticism. 
In particular, his axioms are so strong that they entail 
a \emph{modal collapse} 
\cite{Sobel1987,sobel2004logic}: 
everything that is the case is so necessarily.  
There has been an impressive body of recent and ongoing work
(cf.~\cite{sobel2004logic,Fitting,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,bjordal99,fuhrmann05:_exist_notwen,Hajek2002,Hajek2008,ContemporaryBibliography}
and the references therein) proposing solutions for the modal
collapse.  The goal of this contribution is to discuss the modal
collapse from the point of view of the modal square of opposition.

@@

Independently of the variant of the ontological argument 
under consideration, the following can be said about 
classes of collapsing propositions:
\begin{enumerate}
\item Valid propositions are collapsing: 
if $\phi$ is valid, then $\collapse{\phi}$ is valid.
%
\item The class of collapsing propositions 
is closed under logical equivalence: 
if $\collapse{\phi}$ is valid and $\phi \biimp \phi'$ is valid, 
then $\collapse{\phi'}$ is valid.
%
\item The class of collapsing propositions is 
not generally closed under equi-validity: 
even if $\collapse{\phi}$ is valid and $\phi$ and $\phi'$ 
are equi-valid, \\ $\collapse{\phi'}$ may not be valid.
%
\item The class of collapsing propositions is 
not generally closed under implication: 
even if $\collapse{\phi}$ is valid and 
$\phi \imp \phi'$ is valid, 
$\collapse{\phi'}$ may not be valid.
\end{enumerate}

>>>>

%----------classification, keywords, date
\subjclass{
Prim. 03A02;  % Philosophical aspects of logic and foundations
Sec. 68T02 % Artificial Intelligence
\marginpar{Change?: 03B15, 03B45, 03B35, 68T15}
}

\keywords{Modal Logics, Higher-Order Logics, Ontological Argument,
  Interactive and Automated Theorem Proving}

@@

Nevertheless, G{\"o}del's work still leaves room for criticism.  In
particular, his axioms are so strong that, when assuming unrestricted
comprehension principles\footnote{A possible direction to remedy modal
  collapse, as studied e.g. by Koons
  \cite{koons06:_sobel_goedel_ontol_proof} is to impose restrictions
  on the domain of properties.}, 
\marginpar{added footnote}
they entail a \emph{modal collapse}
\cite{Sobel1987,sobel2004logic}: everything that is the case is so
necessarily.  There has been an impressive body of recent and ongoing
work
(cf.~\cite{sobel2004logic,Fitting,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,bjordal99,fuhrmann05:_exist_notwen,Hajek2002,Hajek2008,ContemporaryBibliography}
and the references therein) proposing solutions for the modal
collapse.  The goal of this contribution is to discuss the modal
collapse from the point of view of the modal square of opposition.
\marginpar{@Bruno: Shall we add a sentence or two here?}

@@

Independently of the variant of the ontological argument 
under consideration, the following can be said about 
classes of collapsing propositions:
\begin{enumerate}
\item Valid propositions are collapsing: 
if $\phi$ is valid, then $\collapse{\phi}$ is valid.
%
\item The class of collapsing propositions 
is closed under logical equivalence: 
if $\collapse{\phi}$ is valid and $\phi \biimp \phi'$ is valid, 
then $\collapse{\phi'}$ is valid.
%
\item The class of collapsing propositions is 
not generally closed under equi-validity: 
even if $\collapse{\phi}$ is valid and $\phi$ and $\phi'$ 
are equi-valid, \\ $\collapse{\phi'}$ may not be valid.
%
\item The class of collapsing propositions is 
not generally closed under implication: 
even if $\collapse{\phi}$ is valid and 
$\phi \imp \phi'$ is valid, 
$\collapse{\phi'}$ may not be valid.
\end{enumerate} \marginpar{@Bruno: Shall we respond here to the
  rewievers criticism?}
